---
title: 'Final Project: ggvis'
author: "Luisa Simpao & Kelly Ye"
date: "March 15, 2019"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to ggvis

The ggvis package allows for data visualization. It allows users to

* Describe data graphics using a similar syntax as the ggplot2 package.
* Produce interactive graphics that can modified locally in RStudio or on a browser.
* Work well with the Shiny platform to publish interactive graphics to use on any browser.

This document allows you to understand the basics of ggvis' tools and how to apply them to data. Because data is often so large, it may be hard to understand and visualization allows us to see patterns and draw conclusions. 

## Data: adminpred

To explore the visualization capabilities of ggvis, we'll use adminpred. This data comes from [Mohan S Acharya](https://www.kaggle.com/mohansacharya/graduate-admissions "Graduate Admissions | Kaggle"). It contains contains Serial No., GRE Scores, TOEFL Scores, University Ratings, Strength of the Statement of Purpose, Strength of the Letter of Recommendation, Undergraduate GPA, Research Experience, and Chance of Admission.

University Ratings, Strength of the Statement of Purpose, and Strength of the Letter of Recommendation are factors on five-point scales. Research Experience is an indicator variable (0 = without, 1 = with).

```{r data_setup, include = FALSE}
library(readr)
adminpred <- read_csv("Admission_Predict.csv")
names(adminpred) <- c("serial","gre","toefl","univ_rate","state_purp","letter_rec","ug_gpa", "research_exp","admit_chance")
adminpred$research_exp <- as.factor(adminpred$research_exp)
adminpred$state_purp <- as.factor(adminpred$state_purp)
adminpred$letter_rec <- as.factor(adminpred$letter_rec)
adminpred$univ_rate <- as.factor(adminpred$univ_rate)
```

```{r load_pkg_data, message = FALSE, collapse = TRUE}
library(tidyverse)
library(ggvis)
library(shiny)

adminpred
```

## Plotting basics

Every ggvis graphic starts by calling ggvis(). The ggvis syntax provides many similarities to those of ggplot2. Where ggplot2 uses:

> data %>% ggplot(aes(x, y, fill)) + geom_FUNCTION_NAME() + layers...

ggvis uses ~variable to mean "by variable":

> data %>% ggvis(~x, ~y) %>% layer_FUNCTION_NAME() %>% more layers...

Notice that for both ggplot2 and ggvis, the data is piped into the function, where the variables are defined. However, they are different in that ggvis continues to pipe the main function to layers, whereas ggplot2 adds layers to the main function.

### Plot options

Let's start by plotting the GRE scores against TOEFL scores.

```{r start}
adminpred %>% ggvis(~gre, ~toefl)
```

ggvis is smart enough to guess what the variables are for! The code above was generated without defining any visual layer, whereas a ggplot() call would also require '+ geom_point()' for this same plot to be created.

### Plots

The ggvis package has many different plot options:

* layer_points() for scatterplots; use an := to customize options
    + size := (a numeric value)
    + shape := (i.e. "diamond", "square")
    + stroke := (a color value)
    + fill := (a color value)
* layer_smooths() for smoothed/regression lines
* layer_model_predictions() for predicting models
    + use model =  to call for a type of model; this needs to be defined for the model to run properly; notice that it is not :=
* layer_boxplots() for box plots

First, an example of a regression line added on to a scatter plot:
```{r reg_plots}
adminpred %>% 
  ggvis(~gre, ~toefl) %>% 
  layer_points(opacity := 0.7) %>% 
  layer_smooths(stroke := "green")
```

Distinct groups can be shown in visualizations, similar to how they are shown with ggplot2. Instead of using a separate group_by() statement, the fill option is used to show the groups. In the example below, shows whether applicants have or have not had research experience. Additionally, a linear model -- as defined by the model option -- is overlay over the data points.
```{r lm_plot}
adminpred %>% 
  ggvis(~gre, ~toefl) %>%
  layer_points(fill = ~factor(research_exp)) %>% # note that it is only an "=" here, rather than ":=" 
  layer_model_predictions(model = "lm")
```

The layer layer_model_predictions() is piped into a group_by() function to create separate linear models for applicants with or without research experience.
```{r grouped}
adminpred %>%
  ggvis(~gre, ~toefl, fill = ~factor(research_exp)) %>%
  layer_points() %>%
  group_by(research_exp) %>%
  layer_model_predictions(model = "lm")
```

A categorical variable is used as the x variable to make the box plots for. For this plot, GRE Scores (gre) are shown on the y-axis; the box plot shows the spread of applicant's GRE scores for each type of university.

```{r boxplot, warning = FALSE}
adminpred %>% 
  ggvis(~factor(univ_rate), ~gre) %>% 
  layer_boxplots()
```

### Graphs

Graphs can also be created using ggvis! Graphs that can be made include:

* layer_bars() for bar graphs
    + use width = to specify the width of the bar
* layer_histograms() for a histogram; note that this only needs an x variable
    + use width = to specify the width of the bar; in this case, a greater width increases the range of x covered by the bar
* layer_lines() fpr a line graph
  
```{r bar_graph}
adminpred %>% 
  ggvis(~gre, ~toefl) %>%
  layer_bars()
```
```{r histogram}
adminpred %>% 
  ggvis(~gre) %>%
  layer_histograms(width = 5)
```
```{r line_graph}
adminpred %>% 
  ggvis(~gre, ~toefl) %>%
  layer_lines()
```

***

## Interactivity

Interactivity in ggvis allows users to customize the look of the graphs they make. ggvis interactivity is built on top of Shiny's reactive programming model, but it is possible to use ggvis without knowing how Shiny works to make simple interactions.

This part describes the basics for a quick and easy way to add basic interactivity to a plot.

A plot with basic interactive controls looks very similar to a static plot, but we replace constant options in functions to produce interactive controls.

For exapmple, input_slider() inside the calls for layer_points() and layer_smooths() allows user to change size of the points or smoothness of the line:
```{r}
adminpred %>%
  ggvis(x = ~gre, y = ~toefl) %>%
  layer_smooths(span = input_slider(0, 1, value = 1)) %>%
  layer_points(size := input_slider(1, 10, value = 10))
```

Interactive plots are built with shiny, so only one can run at a time in an R session. Once finished with a plot, press the stop button in Rstudio or close the browser window and then press Escape or Ctrl + C to stop.

There are a lots of interactive controls we could use for the layers in ggvis, for examples:

* input_slider() produces a slider
* input_checkbox() produces a check-box
* input_checkboxgroup() produces a group of check boxes
* input_numeric() produces a spin box
* input_radiobuttons() produces pick one from a set options
* input_select() produces create a drop-down text box
* input_text() produces arbitrary text input

***

## More Layers

ggvis also has more graphs that what's already been shown and what is outlined here. An example for layer_text() is shown below.

Paths and polygons

* layer_paths() 
    + fill := (a color value); if this option is used, a polygon can be created
* layer_ribbons() fills in the area above the curve (y); the fill can be controlled by defining y and y2
* layer_rects() makes rectangles, defined by x1, y1, x2, and y2
* layer_text() labels points with text when defined with x and y. 

To find the top-scoring applicant in the GRE and TOEFL, layer_text() is used to display Serial No. with the scores. The data is filtered for those without research experience, only to narrow down the data displayed.
```{r text}
adminpred %>% 
  filter(research_exp == 0) %>% 
  ggvis(~gre, ~toefl, text := ~serial, fontSize := 15) %>%
  layer_text()
```

***

## Customize Appearance

Piping the visualizations made to more layers can allow users to add axes and legends, define scales, and customize the appearance of the graphs using properties.

### Axes and legends

To modify the axis' label, when piped into a ggvis function:

> add_axis("x", title = "the x variable")

Axes placement can be controlled, using the orient argument or the offset argument.

> add_axis("x", orient = "top/bottom/left/right")
> add_axis("x, offset = N)

To add a legend, the first line of code may be used below may be piped into a ggvis function. The second one allows for multiple scales to be displayed in one legend. Note that legends are not defined by variables but by arguments in the layer that defines the graph being produced. Examples are shown here:

> add_legend("fill", title = "fill variable")
> add_legend("fill", "size")

To combine legends, one would have to manually do so, by supplying the name of multiple scales to one legend. An example of this will be shown below.

### Scales

In ggvis, scales control the mapping between the data values and values interpreted by the drawing device, just like ggplot2. However, comparing to ggplot2, ggvis has far fewer scales (3 vs 70), and with each function doing much more. 

There are three basic scales in ggvis, which are called the three basic vega scales:

* scale_numeric() for quantitative, numeric values
* scale_ordinal() for qualitative, categorical values
* scale_time() for date/time values

By default, scales are added so that every scaled property gets a scale with the same name as the property (with a few exceptions y2 to y, x2 to x, fillOpacity and strokeOpacity to opacity and so on.).

### Scale arguments
These are some arguments for the scales:

* name: the argument as a string identifier for the scale.
* domain: the input data values to the scale. 
* range: the output visual values from the scale. Usually a vector of length two for quantiative scales, for ordinal scales, it’s a vector the same length as the domain.
* reverse: a boolean flag that will flip the order of the range.

There is a lot we can play with for scales and properties. For example, we could do multiple scales for one property.

### Properties for visual customization

Axes and legends share properties that are defined in a list of props().In the example below, multiple scales are used for one property - two x scales are used to show the different values on the y. Take note of the code below to see how axes, scales, and properties work together.

```{r all_together}

adminpred %>% ggvis(y = ~admit_chance, 
            shape = ~factor(univ_rate)) %>%
  layer_points(prop("x", ~gre, 
            scale = "x_gre"), 
            fill := "pink", 
            stroke := "gray4") %>%
  layer_points(prop("x", ~toefl, 
            scale = "x_toefl"), 
            fill := "lightseagreen", 
            stroke := "gray4") %>%
  add_axis("x", 
           scale = "x_gre", 
           title = "GRE Score", 
           orient = "bottom") %>%
  add_axis("x", 
           scale = "x_toefl", 
           title = "TOEFL Score", 
           orient = "top", 
           properties = axis_props(
             grid = list(stroke = "powderblue"),
             ticks = list(stroke = "powderblue"),
             labels = list(fill = "lightseagreen"))) %>% 
  add_axis("y", 
           title = "Chance of Admissions") %>% 
  add_legend("shape",
           title = "University Rating")
```

Observe below how multiple properties are used for one scale. We force stroke and fill to use the same scale.
```{r}
adminpred %>% 
  ggvis(x = ~gre, 
        y = ~admit_chance, 
        stroke = ~ug_gpa, 
        fill = ~toefl, 
        strokeWidth := 5, 
        size := 200) %>%
  layer_points() %>%
  add_legend("stroke", 
             properties = legend_props(legend = list(y = 50)))
```



