---
title: "Lab 6 Simulation"
author: "Luisa Simpao"
date: "February 26, 2019"
output: html_document
---

```{r load package}
library(tidyverse)
```

# Normal Approximation to the Binomial

1. Generate 1000 Samples from a binomial distribution with 20 trials & a 90% probability of success.
```{r binomial w 20 trials & 90% success}
my_samples <- data.frame(
  binom1 = rbinom(1000, size = 20, prob = 0.90)
)

#a Histogram
ggplot(my_samples, aes(x = binom1)) + geom_histogram()

#b Density plot
ggplot(my_samples, aes(x = binom1)) + geom_density()
```

2. Generate 1000 samples each from a binomial distribution with (a) 90%, (b) 70%, and (c) 50% chance of success.
```{r 1000 samples from binomial w diff success %}
success = c(0.90, 0.70, 0.50)

my_samples2 <- map_dfc(success, ~rbinom(1000, size = 20, prob = .x))

# Plot densities on top of each other
names(my_samples2) <- success

my_samples2 <- my_samples2 %>% 
  gather(key = "Success", value = "Samples")

ggplot(my_samples2, aes(x = Samples, fill = Success)) + 
  geom_density(alpha = 0.5)
```

3. Generate 1000 samples each from a binomial distribution with (a) 20, (b) 50, (c) 100 trials & a 90% chance of success
```{r 1000 samples from binomial w diff # of trials at 90% success}
trials = c(20, 50, 100)

my_samples3 <- map_dfc(trials, ~rbinom(1000, size = .x, prob = 0.90))

names(my_samples3) <- trials

my_samples3 <- my_samples3 %>% 
  gather(key = "Trials", value = "Samples")

my_samples3 <- my_samples3 %>% 
  mutate(Samples = as.numeric(Samples)/as.numeric(Trials))

ggplot(my_samples3, aes(x = Samples, fill = Trials)) + 
  geom_density(alpha = 0.5)

# This is where I make the function! 

plot_prop <- function(trials, prob_success){
  
  my_samplesN <- map_dfc(trials, ~rbinom(1000, size = .x, prob = prob_success))
  
  names(my_samplesN) <- trials
  
  my_samplesN <- my_samplesN %>% 
    gather(key = "Trials", value = "Samples")
  
  my_samplesN <- my_samplesN %>% 
    mutate(Samples = as.numeric(Samples)/as.numeric(Trials))
  
  ggplot(my_samplesN, aes(x = Samples, fill = Trials)) + 
    geom_density(alpha = 0.5)
}

plot_prop(c(20,30,40,50), 0.9)
plot_prop(c(60,70,80,90), 0.9)
```

4. There must be approximately more than 50 trials for normal approximation to be suitable for binomial data. normality becomes more apparent as the number of trials increases.

# The Central Limit Theorem

1. (see below)
```{r normal distribution - mean 50, sdev 10}

# a 1000 samples of size 30
my_samples4 <- replicate(1000, rnorm(30, mean = 50, sd = 10), simplify = FALSE)

# my range function
range <- function(x){
  max(x) - min(x)
}

# 3rd quartile function
quart <- function(x){
quantile(x, probs = 0.75)
}

my_stats <- data.frame(
  means = map_dbl(my_samples4, mean),
  medians = map_dbl(my_samples4, median),
  sdevs = map_dbl(my_samples4, sd), 
  ranges = map_dbl(my_samples4, range),
  quarts = map_dbl(my_samples4,quart)
)

# b Sampling distribution of means
ggplot(my_stats, aes(x = means)) + geom_histogram(bins = 20)

# c Sampling distributions of other statistics in single plotting window
library(ggpubr)

m <- ggplot(my_stats, aes(x = medians)) + geom_histogram(bins = 20)
s <- ggplot(my_stats, aes(x = sdevs)) + geom_histogram(bins = 20)
r <- ggplot(my_stats, aes(x = ranges)) + geom_histogram(bins = 20)
q <- ggplot(my_stats, aes(x = quarts)) + geom_histogram(bins = 20)

ggarrange(m, s, r, q, 
          ncol = 2, nrow = 2)
```

2. (see below)
```{r uniform distribution, min = 10, max = 90}
# a 
my_samples5 <- replicate(1000, runif(30, min = 10, max = 90), simplify = FALSE)

my_stats2 <- data.frame(
  means = map_dbl(my_samples5, mean),
  medians = map_dbl(my_samples5, median),
  sdevs = map_dbl(my_samples5, sd), 
  ranges = map_dbl(my_samples5, range), 
  quarts = map_dbl(my_samples5,quart)
)

# b 
ggplot(my_stats2, aes(x = means)) + geom_histogram(bins = 20)

# c
m <- ggplot(my_stats2, aes(x = medians)) + geom_histogram(bins = 20)
s <- ggplot(my_stats2, aes(x = sdevs)) + geom_histogram(bins = 20)
r <- ggplot(my_stats2, aes(x = ranges)) + geom_histogram(bins = 20)
q <- ggplot(my_stats2, aes(x = quarts)) + geom_histogram(bins = 20)

ggarrange(m, s, r, q, 
          ncol = 2, nrow = 2)
```