---
title: "HipHop Activity"
author: "Luisa Simpao"
date: "1/10/2019"
output: html_document
---

```{r}

# (1)
hiphop <- read.csv("https://raw.githubusercontent.com/kbodwin/STAT-331/master/In-Class%20Activities/Data/hiphop.csv?token=AVHCwRXHTU_HZuxL-Iofhjhi-QqAgfxOks5cSRj-wA%3D%3D")
library(tidyverse)

```

``` {r}

head(hiphop)
str(hiphop)

```
  (2) - using code from Chunk 2 (above)
  There are 38 variables in this data set, encompassing the different types of variables in R.
  Factors: word (64 levels), subj (168 levels), sex (2 levels), ethnic (9 levels)
  Numeric variables: city, county, bieber, asianMove, blackMove, hispanicMove, nativeMove, SAAMEMove, whiteMove, blackWeekly
  Integer variables: age, intl, vocal, classical, folk, rock, country, pop, alternative, hiphop, unclassifiable, countyBlack, jayz, barkley, boondocks, monique, jayz1, boondocks1, familiarity1, trial , fam1. popularityScore, numPreferredArtists, numOverallArtists
  
  (3) - using code from Chunk 2 (above)
  The table 10752 x 38, having 10,752 rows and 38 columns.
  There are 38 columns for the 38 variables, and 10,752 observations for the 64 words being tested on 168 subjects.
  
```{r}

# (4)
# The code below calls a list of the 64 AAE terms
unique(hiphop$word)

```

```{r}
# (5)
missing <- is.na(hiphop)
levels(missing) <- c(TRUE,FALSE)
count_missing <- factor(missing)
summary(missing)
# numPreferredArtists and numOverallArtists have missing values

# (6)
# There are 5632 missing values for the two variables, a total of 11264 missing values.

```

```{r}

# (7)
mean(hiphop$numPreferredArtists, na.rm = TRUE)
sqrt(var(hiphop$numPreferredArtists, na.rm = TRUE))
# The mean is 0.3385 and the standard deviation is 0.7357.

```

```{r}

# 8
subject19 <- hiphop[hiphop$subj == "p19",]
str(subject19)
# The data frame subject19 has 64 rows and 38 columns. 

```

```{r}

# (9)
# Below is code to display the values for the familiarity variable for subject19
subject19$familiarity
# Below is code to obtain simple statistcs about the familiarity variable for subject19
summary(subject19$familiarity)

```

```{r}

# 10
subject19 <- subject19[order(subject19$familiarity),]

```